---
title: "Confirmatory factor analysis of MASQ anhedonia, GAD-7 and PHQ-9 in RAMP baseline data"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

This workbook will conduct the factor analysis of MASQ, GAD and PHQ-9 data in RAMP baseline to establish whether it corresponds to a three factor model, corresponding to the tripartite theory. 


#Set up 

```{r Delete everything in your global environment}
remove(list = ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment=NA,
                      prompt=FALSE,
                      cache=FALSE)
```

Note: load tidyverse last!
```{r setup, Load packages}
if(!require(summarytools)){
  install.packages("summarytools")
  library(summarytools)
}
if(!require(ggplot2)){
  install.packages("ggplot2")
  library(ggplot2)
}

if(!require(psych)){
  install.packages("psych")
  library(psych)
}

if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}


if(!require(skimr)){
  install.packages("skimr")
  library(skimr)
}

if(!require(lavaan)){
  install.packages("lavaan")
  library(lavaan)
}

if(!require(geom_text)){
  install.packages("geom_text")
  library(geom_text)
}

```

```{r Set GLAD palette for plotting}
glad_palette = c("#efc00b", "#b7dee8")
```

# Read in data

```{r Read in GAD, PHQ and MASQ data from the RAMP baseline cohort}
# Read in the data
MASQ <- readRDS(file =  "../../../data_clean/masq/masq.clean_baseline_items.RAMP_notrecoded.RDS")
GAD7 <- readRDS(file =  "../../../data_clean/gad/gad.clean_baseline_items.RAMP.RDS")
PHQ9 <- readRDS(file =  "../../../data_clean/phq/phq.clean_baseline_items.RAMP.RDS")

```

# reverse MASQ scores to be in line with GAD & PHQ 

also make -77 NA
```{r recode MASQ}

MASQ.recode <- MASQ %>%
  mutate_at(vars(starts_with("masq")),
            funs(case_when(
             . == 1 ~ 5,
             . == 2 ~ 4,
             . == 3 ~ 3,
             . == 4 ~ 2,
             . == 5 ~ 1,
             . == -77 ~ NA_real_,
             TRUE ~  NA_real_)))
  
  
```


# merge MASQ, PHQ and GAD data

```{r merge data}

merged_data <- left_join(GAD7,PHQ9)
merged_data <- left_join(merged_data,MASQ.recode)

```

## check data names and dimensions

```{r check data}

head(merged_data)
names(merged_data)
dim(merged_data)
```

# drop uneccessary columns, blanks and duplicates

```{r dropscols and blanks}

cfa_data.clean <- merged_data %>% #new dataset with ID
  drop_na(ID) %>% # Drop NAs
  select(-c(sample,startDate,endDate))

names(cfa_data.clean)

```



```{r drop duplicates}

##Identify dup IDs. These seem to be cases where we randomly dropped different people from different datasets due to actually the same IDs occurring
cfa_data.clean  %>%
   group_by(ID) %>%
   summarize(N = n()) %>%
   filter(N > 1)

## randomly retain 1 of each duplicate - whichever happens to be last

cfa_data.clean <- cfa_data.clean %>% 
    group_by(ID) %>% 
    filter(!duplicated(ID,
                      fromLast = TRUE)
           ) %>%
  ungroup()

```

# Correlations

check the correlations between all items

```{r upper and lower triangle functions}
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
```

```{r check item correlation}

cormat <- round(cor(!is.na(cfa_data.clean[,2:27])),2)

upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)

# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()



ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))


print(ggheatmap)
```

All basically very highly correlated. check lowest correlation between two item pairs:

```{r check lowed cor}

min(melted_cormat$value)

melted_cormat[(melted_cormat$value == 0.63),]
```
# CFA

## a prior model specficiation
What items will work best for the tripartite: specify the model

***Item pool***
gad7.pandemic_nervous_anxious_on_edge
gad7.pandemic_cant_control_worry
gad7.pandemic_worry_too_much
gad7.pandemic_trouble_relaxing
gad7.pandemic_restless
gad7.pandemic_easily_annoyed
gad7.pandemic_afraid_something_terrible_will_happen
phq9.pandemic_little_interest_or_pleasure
phq9.pandemic_down_depressed_hopeless
phq9.pandemic_trouble_falling_or_staying_asleep
phq9.pandemic_feel_tired_little_energy
phq9.pandemic_poor_appetite_or_overeating
phq9.pandemic_feeling_like_a_failure
phq9.pandemic_trouble_concentrating
phq9.pandemic_fidgeting_or_moving_slowly
phq9.pandemic_thoughts_better_off_dead
masq.pandemic_felt_successful
masq.pandemic_felt_really_happy
masq.pandemic_felt_optimistic
masq.pandemic_lot_felt_fun
masq.pandemic_felt_like_i_accomplished_a_lot
masq.pandemic_forward_lot_felt
masq.pandemic_felt_really_talkative
masq.pandemic_felt_really_up_or_lively
masq.pandemic_lot_felt_energy
masq.pandemic_felt_really_good_about_myself


**Kirstin item to factor spec:**

*General distress*
gad7.pandemic_nervous_anxious_on_edge
gad7.pandemic_cant_control_worry
gad7.pandemic_worry_too_much
gad7.pandemic_easily_annoyed
gad7.pandemic_afraid_something_terrible_will_happen
phq9.pandemic_down_depressed_hopeless
phq9.pandemic_trouble_falling_or_staying_asleep
phq9.pandemic_feel_tired_little_energy
phq9.pandemic_poor_appetite_or_overeating
phq9.pandemic_feeling_like_a_failure
phq9.pandemic_trouble_concentrating
phq9.pandemic_thoughts_better_off_dead

*Negative affect (anhedonia)*
phq9.pandemic_little_interest_or_pleasure
masq.pandemic_felt_successful
masq.pandemic_felt_really_happy
masq.pandemic_felt_optimistic
masq.pandemic_lot_felt_fun
masq.pandemic_felt_like_i_accomplished_a_lot
masq.pandemic_forward_lot_felt
masq.pandemic_felt_really_talkative
masq.pandemic_felt_really_up_or_lively
masq.pandemic_lot_felt_energy
masq.pandemic_felt_really_good_about_myself

*physiological arousal*
gad7.pandemic_restless
gad7.pandemic_trouble_relaxing
phq9.pandemic_fidgeting_or_moving_slowly


**Katie item to factor spec**
*General distress*
gad7.pandemic_cant_control_worry
gad7.pandemic_worry_too_much
gad7.pandemic_easily_annoyed
gad7.pandemic_afraid_something_terrible_will_happen
phq9.pandemic_down_depressed_hopeless
phq9.pandemic_trouble_falling_or_staying_asleep
phq9.pandemic_feel_tired_little_energy
phq9.pandemic_poor_appetite_or_overeating
phq9.pandemic_feeling_like_a_failure
phq9.pandemic_trouble_concentrating
phq9.pandemic_thoughts_better_off_dead
phq9.pandemic_fidgeting_or_moving_slowly

*Negative affect (anhedonia)*
phq9.pandemic_little_interest_or_pleasure
masq.pandemic_felt_successful
masq.pandemic_felt_really_happy
masq.pandemic_felt_optimistic
masq.pandemic_lot_felt_fun
masq.pandemic_felt_like_i_accomplished_a_lot
masq.pandemic_forward_lot_felt
masq.pandemic_felt_really_talkative
masq.pandemic_felt_really_up_or_lively
masq.pandemic_lot_felt_energy
masq.pandemic_felt_really_good_about_myself

*physiological arousal*
gad7.pandemic_restless
gad7.pandemic_trouble_relaxing
gad7.pandemic_nervous_anxious_on_edge


***Differences: (1) nervous, anxious and on edge (Katie Phys arousal, Kirstin general distress), (2) fidgeting or moving slowly (Katie general distress, Kirstin Physiological arousal)***

**Final item to factor spec after consulting on any differences**
(still being confirmed. Consulting with others and reading papers to align as closely as possible with theory)

*General distress*
gad7.pandemic_cant_control_worry
gad7.pandemic_worry_too_much
gad7.pandemic_easily_annoyed
gad7.pandemic_afraid_something_terrible_will_happen
phq9.pandemic_down_depressed_hopeless
phq9.pandemic_trouble_falling_or_staying_asleep
phq9.pandemic_feel_tired_little_energy
phq9.pandemic_poor_appetite_or_overeating
phq9.pandemic_feeling_like_a_failure
phq9.pandemic_trouble_concentrating
phq9.pandemic_thoughts_better_off_dead


*Negative affect (anhedonia)*
phq9.pandemic_little_interest_or_pleasure
masq.pandemic_felt_successful
masq.pandemic_felt_really_happy
masq.pandemic_felt_optimistic
masq.pandemic_lot_felt_fun
masq.pandemic_felt_like_i_accomplished_a_lot
masq.pandemic_forward_lot_felt
masq.pandemic_felt_really_talkative
masq.pandemic_felt_really_up_or_lively
masq.pandemic_lot_felt_energy
masq.pandemic_felt_really_good_about_myself

*physiological arousal*
gad7.pandemic_restless
gad7.pandemic_trouble_relaxing
gad7.pandemic_nervous_anxious_on_edge




## specify the CFA in lavaan


```{r specify lavaan model}

tripartite_model <- 'General_distress =~ gad7.pandemic_cant_control_worry + gad7.pandemic_worry_too_much + gad7.pandemic_easily_annoyed + gad7.pandemic_afraid_something_terrible_will_happen + phq9.pandemic_down_depressed_hopeless + phq9.pandemic_trouble_falling_or_staying_asleep + phq9.pandemic_feel_tired_little_energy + phq9.pandemic_poor_appetite_or_overeating + phq9.pandemic_feeling_like_a_failure + phq9.pandemic_trouble_concentrating + phq9.pandemic_thoughts_better_off_dead

negative_affect =~ masq.pandemic_felt_successful + masq.pandemic_felt_really_happy + masq.pandemic_felt_optimistic + masq.pandemic_lot_felt_fun + masq.pandemic_felt_like_i_accomplished_a_lot + masq.pandemic_forward_lot_felt + masq.pandemic_felt_really_talkative + masq.pandemic_felt_really_up_or_lively + masq.pandemic_lot_felt_energy + masq.pandemic_felt_really_good_about_myself + phq9.pandemic_little_interest_or_pleasure

anxious_arousal =~  + gad7.pandemic_restless + gad7.pandemic_trouble_relaxing + phq9.pandemic_fidgeting_or_moving_slowly + gad7.pandemic_nervous_anxious_on_edge'

```

```{r fit cfa}

#fit <- cfa(tripartite_model, data=cfa_data.clean)

```


```{r summarise cfa model}

#summary(fit,fit.measures = TRUE,standardized=TRUE)

```