---
title: "Run Growth Mixture Modelling for RAMP & COPING longitudinal analyses of common mental health symptoms: PHQ"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: hide

html_notebook:
  theme: cerulean
toc: yes
---

# Introduction

## background
This workbook centralises analyses of the PHQ-9 total scores longitudinal analyses in the RAMP and COPING samples - identifying sub classes in overall trajectories of depression symptoms over the first year of the COVID-19 pandemic (Growth Mixture Modelling; GMM). This corresponds with analytic step 2 in the pre-registration for project ["Moderators of symptom trajectories of depression and anxiety during the COVID-19 pandemic"](https://osf.io/jvca5/).

The exploratory approach attempts to decompose heterogeneous trajectories into distinct subclasses. 

To allow estimation of variance while minimising computational requirements, we will constrain the variance of the intercept to be equal across classes, whilst fixing the variance in the slope to zero.

Best fit model from the 1 class lgcm performed (relating to step 1 in the pre-registration) was a piecewise model, with each inflection point dictated by an imposition or easing of lockdown restrictions, allowing for residuals to correlate between contiguous time points.


## Analytic steps

**global MPlus** 
All models will be run using Robust Full Information Maximum Likelihood estimation (MLR in MPlus) to account for missing data (assumed missing at random) robust to skewing of the data. 

In order to avoid a global maxima/local solution problem (log likelihood depends on starting values used, thus could have a false maximum or sub-optimal model) we will use multiple random sets of starting values, and ensure the highest log likelihood is replicated at least twice. In practice, we will use 1000 random sets of starting values, and carry forward the 250 values with the highest log likelihood to final stage optimisations. This will be set using STARTS = 1000 250 in Mplus input scripts.  This should ensure global maxima are achieved, whilst balancing computational requirements. 

**Step 1. Test a k + 1 class model, starting with 2, of the best fit latent growth curve model established in step 1 (lgcm) until any of the following conditions are met:**   

* Indices of model fit show a deterioration in fit according to the BIC relative to a model with fewer classes. The first model (2 class) will be compared to the lgcm (1 class), and so forth.
* The model no longer converges due to complexity   
* We exceed 7 classes (this is the maximum number of classes that would be practically useful and logically interpretable). 
* If the BIC continues to decrease up until 7 classes, then the best model will be ascertained by first examining the proportion of the sample that has been allocated to the smallest trajectory group. If this is < 1%, then the previous class solution will be preferred. 
* If BIC decreases up until 7 class solution, and all classes constrain > 1% of the data, then the best model will be selected by creating a screen plot of BIC values. the model that falls at the inflection point (or the point of diminishing gains) will be selected (see pre registration for citations).
* if inflection point remains ambiguous, then remaining fit indices will be compared globally to select between two similar models. The model with the highest entropy will be selected preferentially. If entropy is equivalent, then model which has better fit on the most number of fit indices (described below) will be selected. 

**Step 2. Assess overall fit of the selected model to ensure it is the best and most parsimonious solution**
* We will ensure there is no evidence for the local maxima problem by rerunning the analysis using he top two seeds from the replicated best log likelihood from the original analysis using OPTSEED in a new Mplus input file. We will check that the best log likelihood from this output matches that from the previous output. We will include TECH11 output in this MPLUS run to output the Vuong-Lo-Mendell-Rubin Likelihood ratios test comparing the best fit model to the earlier fit model

* We will run a second analysis using OPTSEED to be even more ure there is no evidence for the local maxima problem by rerunning the analysis using the top two seeds from the replicated best log likelihood from the original analysis in a new Mplus input file. We will check that the best log likelihood from this output matches that from the previous output. We will include TECH14 output in this MPLUS run to output the Bootstrapped Likelihood ratios test (BLRT) comparing the best fit model to the earlier nested model.

This step wise procedure should save time whilst ensuring we select the optimal model and obtain unbiased estimates. [see this webnote for a description of the step wise procedure](https://www.statmodel.com/examples/webnotes/webnote14.pdf).

* assuming model selection was unambiguous, assess model fit indices globally to ensure all indicators are reasonable   
* assuming above is true, check smallest group has >1% of the sample
* assess parsimony of the model. Does it divide one trajectory into two which only differ in intercept?
* Assess entropy. If less than <0.8, ensure appropriate discussion of this limitation of class assignment. Consider impact of accuracy of class assignment on any subsequent predictive steps 

## fit indices that will be used to identify the best model
See pre registration for detailed description of all outcomes.

* Bayesian Information Criterion (BIC). Decisions of best model will be primarily based on this. Smaller value = better fit
* Aikaike Information Criterion (AIC). Smaller value = better fit
* Likelihood ratio test (LRT) P > 0.05 indicates better model

# Setup

```{r}
# clear global environment
remove(list = ls())
```

***note: some of the install and update processes require command line verification in the console. If you are running this script for the first time and it does not proceed, open and check the console***

```{r setup, include=FALSE}

# Install / load packages

if(!require(knitr)){
  install.packages("knitr")
  library(knitr)
}

if(!require(summarytools)){
  install.packages("summarytools")
  library(summarytools)
}

if(!require(data.table)){
  install.packages("data.table")
  library(data.table)
}


if(!require(kableExtra)){
  install.packages("kableExtra")
  library(kableExtra)
}

if(!require(glue)){
  install.packages("glue")
  library(glue)
}

if(!require(MplusAutomation)){
  install.packages("MplusAutomation")
  library(MplusAutomation)
}



if(!require(texreg)){
  install.packages("texreg")
  library(texreg)
}

if(!require(relimp)){
  install.packages("relimp")
  library(relimp)
}

if(!require(BiocManager)){
  install.packages("BiocManager")
  library(BiocManager)
}

if(!require(rhdf5)){
  BiocManager::install("rhdf5")
  library(rhdf5)
}

if(!require(plyr)){
  install.packages("plyr")
  library(plyr)
}


if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}


if(!require(kableExtra)){
  install.packages("kableExtra")
  library(kableExtra)
}


if(!require(lmtest)){
  install.packages("lmtest")
  library(lmtest)
}




## global workbook settings


knitr::opts_chunk$set(results = 'asis',      
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE, 
                      warning = FALSE,
                      echo = TRUE,
                      fig.height = 5,
                      fig.width = 10,
                      fig.align = "center")

st_options(plain.ascii = FALSE,       
            style = "rmarkdown",      
            footnote = NA,            
            subtitle.emphasis = FALSE)                         


options(scipen = 999)
```


***colour paletttes and plot label list***
RAMP colours (first) from website and second is a generated left tetradic (rectangular) colour complement of the RAMP logo green for contrast and complement. 
```{r colour palette, include=FALSE}

pres.palette <- c("#4ED1B7", "#F6C66A", "#D14D68", "#1A3E72",  "#D1754D","#FFED04","#3F9D8A","#B74DD1") 

prev.pres.palette<-c("#4590B8", "#4D68D1", "#D14D68", "#D1B74D", "#68D14D", "#B74DD1", "#D1754D")

timepoint_list <- c("7th April 2020","19th May 2020","2nd June 2020","16th June 2020","30th June 2020",
                    "14th July 2020", "28th July 2020","25th August 2020","22nd September 2020","20th October 2020",
                    "17th November 2020","15th December 2020","12th January 2021","9th February 2021",
                    "9th March 2021", "6th April 2021")

timepoint_list_anh <- c("19th May 2020","2nd June 2020","16th June 2020","30th June 2020",
                    "14th July 2020", "28th July 2020","25th August 2020","22nd September 2020","20th October 2020",
                    "17th November 2020","15th December 2020","12th January 2021","9th February 2021",
                    "9th March 2021", "6th April 2021")

```


# Step 1: run a K+1 class analysis {.tabset}

Will run each class consecutively, to check model for clear indications of stopping points defined above. 

## Two classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run two class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_2_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never"
          )
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model


##### call in data from previous model

```{r read depression model}

depression.one.class.out <- readModels("./MPlus_input_lgcm/PHQ/phq_latentGrowthCurve_piecewise_wPairwise_correlations.out", 
                           recursive=FALSE)

```



##### call in data from current model

```{r read depression model two}

depression.two.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_2_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from one class}
justBIC_one_class <- depression.one.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

```{r extract bic from two class}
justBIC_two_class <- depression.two.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare one vs two}

BIC_one_two_compare <- rbind(justBIC_one_class,
                         justBIC_two_class)

BIC_one_two_compare
```

Difference in BIC
```{r compare difference one vs two}

BIC_one_two_compare$BIC[1] - BIC_one_two_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference one vs two adjusted}

BIC_one_two_compare$aBIC[1] - BIC_one_two_compare$aBIC[2]
```

## Three classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run three class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_3_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never")
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model

##### call in data from current model

```{r read depression model three}

depression.three.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_3_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from three class}
justBIC_three_class <- depression.three.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare two vs three}

BIC_two_three_compare <- rbind(justBIC_two_class,
                         justBIC_three_class)
BIC_two_three_compare
```

Difference in BIC
```{r compare difference two vs three}

BIC_two_three_compare$BIC[1] - BIC_two_three_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference two vs three adjusted}

BIC_two_three_compare$aBIC[1] - BIC_two_three_compare$aBIC[2]
```


## four classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run four class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_4_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never")
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model

##### call in data from current model

```{r read depression model four}

depression.four.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_4_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from four class}
justBIC_four_class <- depression.four.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare three vs four}

BIC_three_four_compare <- rbind(justBIC_three_class,
                         justBIC_four_class)

BIC_three_four_compare
```

Difference in BIC
```{r compare difference three vs four}

BIC_three_four_compare$BIC[1] - BIC_three_four_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference three vs four adjusted}

BIC_three_four_compare$aBIC[1] - BIC_three_four_compare$aBIC[2]
```


## five classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run five class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_5_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never")
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model

##### call in data from current model

```{r read depression model five}

depression.five.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_5_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from five class}
justBIC_five_class <- depression.five.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare four vs five}

BIC_four_five_compare <- rbind(justBIC_four_class,
                         justBIC_five_class)

BIC_four_five_compare
```

Difference in BIC
```{r compare difference four vs five}

BIC_four_five_compare$BIC[1] - BIC_four_five_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference four vs five adjusted}

BIC_four_five_compare$aBIC[1] - BIC_four_five_compare$aBIC[2]
```


## six classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run six class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_6_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never")
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model

##### call in data from current model

```{r read depression model six}

depression.six.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_6_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from six class}
justBIC_six_class <- depression.six.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare five vs six}

BIC_five_six_compare <- rbind(justBIC_five_class,
                         justBIC_six_class)

BIC_five_six_compare
```

Difference in BIC
```{r compare difference five vs six}

BIC_five_six_compare$BIC[1] - BIC_five_six_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference five vs six adjusted}

BIC_five_six_compare$aBIC[1] - BIC_five_six_compare$aBIC[2]
```

## seven classes

### run model
note, make sure eval=false in this block in final version to ensure script can be knitted even on a computer without mplus.

```{r run seven class model, eval=TRUE}
runModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_7_classes.inp", 
          recursive = FALSE,
          showOutput=TRUE,
          replaceOutfile="never")
```


### Evaluate model {.tabset}

#### Compare BIC of current and previous model

##### call in data from current model

```{r read depression model seven}

depression.seven.class.out <- readModels("./MPlus_input_gmm/PHQ/phq_GMM_piecewise_7_classes.out", 
                           recursive=FALSE)

```


##### extract BIC 

```{r extract bic from seven class}
justBIC_seven_class <- depression.seven.class.out$summaries  %>%
  select(c("Title",  "BIC", "aBIC")) %>%
  arrange(BIC)

```

##### compare

```{r create BIC table and compare six vs seven}

BIC_six_seven_compare <- rbind(justBIC_six_class,
                         justBIC_seven_class)

BIC_six_seven_compare
```

Difference in BIC
```{r compare difference six vs seven}

BIC_six_seven_compare$BIC[1] - BIC_six_seven_compare$BIC[2]
```

Difference in sample size adjusted BIC
```{r compare difference six vs seven adjusted}

BIC_six_seven_compare$aBIC[1] - BIC_six_seven_compare$aBIC[2]
```



# Step 2: Assess overall fit of the selected model

## Check for local maxima problem (replicate best log likelihood) and perform Likelihood ratio tests

run optseed with TECH 11
run optseed again with TECH 14



# Supplementary table of fit indices for all models run

read all GMM models
```{r read all models} 

depression.all.out <- readModels("./MPlus_input_gmm/PHQ", 
                           recursive=FALSE)

```

extract relevant summaries from GMM and 1 class lgcm

```{r extract gmm summaries}
justSummariesGMM <- do.call("rbind.fill",
  sapply(depression.all.out ,"[", "summaries")) %>%
  select(c("Title", "Observations", "Parameters","NGroups", "AIC", "BIC", "LL")) %>%
  arrange(BIC)
```

extract the same from the one trajectory lgcm

```{r extract all sums from one trajectory}

justSummarieslgcm <- do.call(sapply(depression.one.class.out ,"[", "summaries")) %>%
  select(c("Title", "Observations", "Parameters","NGroups","AIC", "BIC",  "LL"))

```

And bind these together to print in a nice table

```{r build pretty summary table}

justSummaries <- rbind(justSummariesGMM,justSummarieslgcm) %>%
  arrange(BIC)

justSummaries %>%
  kbl() %>%
  kable_styling()

```





### add LRT output these as a column to our table and print again

# Plots {.tabset}

## One classes

```{r Format latent growth curves data, include=FALSE}

dat_plot_curves_1<-data.frame(matrix(nrow=16, ncol=2))

colnames(dat_plot_curves_1)<-c("Group","Observed")

dat_plot_curves_1$Group<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values
dat_plot_curves_1$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_1$timepoint<-c(0:15) 

dat_plot_curves_1<-dat_plot_curves_1%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Group","Observed")))

dat_plot_curves_1 <- dat_plot_curves_1 %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms One class}

growth_curves_oneclass <- ggplot() +
  geom_line(data = dat_plot_curves_1,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_1,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(6,10), breaks=seq(6, 10, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_oneclass



```

## Two classes

```{r Format latent growth curves data two class, include=FALSE}

dat_plot_curves_two<-data.frame(matrix(nrow=16, ncol=3))

colnames(dat_plot_curves_two)<-c("Class1","Class2","Observed")

dat_plot_curves_two$Class1<- depression.two.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_two$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_two$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_two$timepoint<-c(0:15) 

dat_plot_curves_two<-dat_plot_curves_two%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Observed")))

dat_plot_curves_two <- dat_plot_curves_two %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms two class}

growth_curves_twoclass <- ggplot() +
  geom_line(data = dat_plot_curves_two,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_two,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(4,9), breaks=seq(4,9, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_twoclass



```


## three classes

```{r Format latent growth curves data three class, include=FALSE}

dat_plot_curves_three<-data.frame(matrix(nrow=16, ncol=4))

colnames(dat_plot_curves_three)<-c("Class1","Class2","Class3","Observed")

dat_plot_curves_three$Class1<- depression.three.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_three$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_three$Class3<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[3]
dat_plot_curves_three$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_three$timepoint<-c(0:15) 

dat_plot_curves_three<-dat_plot_curves_three%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Class3", "Observed")))

dat_plot_curves_three <- dat_plot_curves_three %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms three class}

growth_curves_threeclass <- ggplot() +
  geom_line(data = dat_plot_curves_three,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_three,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(7,16), breaks=seq(7,16, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_threeclass

```

## four classes

```{r Format latent growth curves data four class, include=FALSE}

dat_plot_curves_four<-data.frame(matrix(nrow=16, ncol=5))

colnames(dat_plot_curves_four)<-c("Class1","Class2","Class3","Class4","Observed")

dat_plot_curves_four$Class1<- depression.four.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_four$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_four$Class3<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[3]
dat_plot_curves_four$Class4<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[4]
dat_plot_curves_four$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_four$timepoint<-c(0:15) 

dat_plot_curves_four<-dat_plot_curves_four%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Class3","Class4", "Observed")))

dat_plot_curves_four <- dat_plot_curves_four %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms four class}

growth_curves_fourclass <- ggplot() +
  geom_line(data = dat_plot_curves_four,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_four,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(7,24), breaks=seq(7,24, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_fourclass

```

## five classes

```{r Format latent growth curves data five class, include=FALSE}

dat_plot_curves_five<-data.frame(matrix(nrow=16, ncol=6))

colnames(dat_plot_curves_five)<-c("Class1","Class2","Class3","Class4","Class5","Observed")

dat_plot_curves_five$Class1<- depression.five.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_five$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_five$Class3<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[3]
dat_plot_curves_five$Class4<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[4]
dat_plot_curves_five$Class5<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[5]
dat_plot_curves_five$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_five$timepoint<-c(0:15) 

dat_plot_curves_five<-dat_plot_curves_five%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Class3","Class4","Class5", "Observed")))

dat_plot_curves_five <- dat_plot_curves_five %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms five class}

growth_curves_fiveclass <- ggplot() +
  geom_line(data = dat_plot_curves_five,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_five,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(0,27), breaks=seq(0,27, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_fiveclass

```

## six classes

```{r Format latent growth curves data six class, include=FALSE}

dat_plot_curves_six<-data.frame(matrix(nrow=16, ncol=7))

colnames(dat_plot_curves_six)<-c("Class1","Class2","Class3","Class4","Class5","Class6", "Observed")

dat_plot_curves_six$Class1<- depression.six.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_six$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_six$Class3<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[3]
dat_plot_curves_six$Class4<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[4]
dat_plot_curves_six$Class5<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[5]
dat_plot_curves_six$Class6<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[6]
dat_plot_curves_six$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_six$timepoint<-c(0:15) 

dat_plot_curves_six<-dat_plot_curves_six%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Class3","Class4","Class5","Class6","Observed")))

dat_plot_curves_six <- dat_plot_curves_six %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms six class}

growth_curves_sixclass <- ggplot() +
  geom_line(data = dat_plot_curves_six,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_six,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(0,27), breaks=seq(0,27, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_sixclass

```
## seven classes

```{r Format latent growth curves data seven class, include=FALSE}

dat_plot_curves_seven<-data.frame(matrix(nrow=16, ncol=8))

colnames(dat_plot_curves_seven)<-c("Class1","Class2","Class3","Class4","Class5","Class6", "Class7", "Observed")

dat_plot_curves_seven$Class1<- depression.seven.class.out$gh5$means_and_variances_data$y_estimated_means$values[1] 
dat_plot_curves_seven$Class2<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[2]
dat_plot_curves_seven$Class3<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[3]
dat_plot_curves_seven$Class4<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[4]
dat_plot_curves_seven$Class5<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[5]
dat_plot_curves_seven$Class6<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[6]
dat_plot_curves_seven$Class7<- depression.one.class.out$gh5$means_and_variances_data$y_estimated_means$values[7]
dat_plot_curves_seven$Observed<- depression.one.class.out$gh5$means_and_variances_data$y_observed_means$values

dat_plot_curves_seven$timepoint<-c(0:15) 

dat_plot_curves_seven<-dat_plot_curves_seven%>%pivot_longer(-timepoint, names_to = "Model", values_to = "score")%>%mutate(Model = factor(Model, levels = c("Class1","Class2","Class3","Class4","Class5","Class6","Class7", "Observed")))

dat_plot_curves_seven <- dat_plot_curves_seven %>%
  mutate(Type = case_when(Model == "Observed"~ "Observed",
                          TRUE ~ "Estimated"))
```



```{r Growth Curve Forms seven class}

growth_curves_sevenclass <- ggplot() +
  geom_line(data = dat_plot_curves_seven,
        aes(x = timepoint, y = score, colour = Model,linetype=Type)) +
  geom_point(data = dat_plot_curves_seven,
        aes(x = timepoint, y = score, colour = Model, shape=Model), show.legend = F) +
  scale_color_manual(values = pres.palette) +
  labs(x = "Follow up time point", y ="PHQ9 Score") +
  scale_y_continuous(expand = c(0,0), limits = c(0,27), breaks=seq(0,27, 1)) + 
  scale_x_continuous(expand = c(0,0), limits = c(-0.5,15.5), breaks=seq(0, 15, by = 1),
                   labels = timepoint_list) +
  theme(panel.grid.major.y = element_line(size = 0.5,
                                        linetype = 'dashed',
                                        colour = "gray"),
        axis.text = element_text(colour="black", size = 14),
        axis.title = element_text(colour="black", size = 16),
        axis.text.x = element_text(angle = 90,
        hjust=1),
        legend.key = element_blank(),
        legend.text = element_text(colour = "black", size = 11),
        panel.background = element_blank()) 

growth_curves_sevenclass

```
