---
title: "Clean baseline predictors for use in multionomial logistic regression"
author: "K L Purves"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: cerulean
code_folding: show

html_notebook:
  theme: cerulean
toc: yes
---

# Summary

Data cleaning for baseline predictors to include in multinomial regressions. The predictors included are those that were significant predictors of change in symptoms from pre-pandemic to pandemic baseline from the first RAMP and COPING paper [see preprint here](https://psyarxiv.com/sf7b6/), and ethnicity, as this has been identified as a risk factor for a range of negative pandemic related outcomes.

This list of predictors has been selected to specifically test the hypothesis that factors that predict a worsening of symptoms during the pandemic may also predict symptom trajectory over the course of the pandemic.

Predictors to include are:    

* Age (categorical)    
* Gender     
* Employment status  (before the pandemic)  
* Pre-existing mental health diagnosis     
* ethnicity^*     


^* We will use the ethnicity variable in two different ways in order to ensure we are making the best possible use of the data we have available for this important predictor. First, we will use each self reported ethnicity category as an individual predictor. Although each individual group may be too small to be well powered to detect small effects, we feel it is important to recognise that experiences of different minoritised ethnic communities are not homogeneous, and these different experiences may have an important impact on the overall trajectory of mental healthy symptoms during the pandemic.By collapsing across different minoritsed ethnic community groups, we may miss important differences in potential risk factors and outcomes associated with any one specific ethnicity or background experience. 

We will then perform a second analysis, where we group all minoritsed ethnic communities together, comparing them to the visible majority white ethnicity group. Whist this grouping does not respect the different experiences that different communities may experience, it does consider the impact of being a "visible minority" and the potential risk for discrimination that this may bring. This analysis will be better powered to detect smaller effects.

We hope that by performing this analysis in both ways, we will be able to fully consider the possible impact of experiences of all ethnic communities on anxiety, depression and anhedonia trajectories throughout the pandemic. 

# Setup

```{r}
# clear global environment
remove(list = ls())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment=NA,
                      prompt=FALSE,
                      cache=FALSE)
```

Note: load tidyverse last!
```{r Load packages}



if(!require(psych)){
  install.packages("psych")
  library(psych)
}

if(!require(kableExtra)){
  install.packages("kableExtra")
  library(kableExtra)
}

if(!require(dplyr)){
  install.packages("dplyr")
  library(dplyr)
}

if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}
```

```{r}
# read in data files
source("../../../Credentials/paths_baseline_cleaning.R")
source("../../../Credentials/paths_anhedonia_analysis.R")
```

## read in all functions from function library

```{r call in function library functions}
# source all functions in the function library folder
files.sources = paste0("../functions/",list.files("../functions"))
sapply(files.sources, source)

```

# COPING data preprocessing {.tabset}

## read in and rename data

```{r read in files}
#COPING
fourcohort <- readRDS(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/full_cohort.rds")

#RAMP - to combine
ramp_age <-readRDS(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/ramp/demographics/age_ramp_clean.rds")

ramp_age <- ramp_age[c(1,2,5)]

ramp_gender <-readRDS(file =
"/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/ramp/demographics/sex_gender_sexuality_ramp_clean.rds")
ramp_gender <- ramp_gender[c(1,6)]

ramp_ethnicity <-readRDS(file =
"/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/ramp/demographics/ethnicity_ramp_clean.rds")
ramp_ethnicity <- ramp_ethnicity[c(1,5)]

ramp_dem <- left_join(ramp_age, ramp_gender, by = "ID")
ramp_dem <- left_join(ramp_dem, ramp_ethnicity, by = "ID")

####

ramp_employment <- readRDS(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/ramp/employ_ramp.rds")
ramp_employment <- ramp_employment[c(4,6:31,50,51)]

names(ramp_employment)[1] <- "ID"

ramp_employment <- ramp_employment %>% mutate(employment_prior_covid = case_when(
  employ.fulltime_employed == 1 | employ.parttime_employed == 1 | employ.zerohours_contract == 1 | employ.selfemployed == 1 | employ.contract_or_freelance_work == 1 | employ.small_business_owner ==1 ~ "Employed",
  employ.unemployed == 1 | employ.stayathome_parent_or_carer == 1 ~ "Unemployed",
  employ.student_ == 1 | employ.student_.1 == 1 ~ "Student",
  employ.retired == 1 ~ "Retired" ))

table(ramp_employment$employment_prior_covid)

ramp_employment <- ramp_employment %>% mutate(employment_change = case_when(
  employ.became_unemployed == 1 | employ.reduction_in_hours == 1 ~ "Decreased employment",
  employ.became_employed == 1 | employ.increased_hours == 1 ~ "Increased employment",
  employ.furloughed_or_paid_leave_ == 1 | employ.furloughed_or_paid_leave_.1 == 1 | employ.paid_leave_furloughed == 1 ~ "Furloughed",
  employ.my_employment_status_has_not_changed == 1 ~"Employment not changed"))

table(ramp_employment$employment_change)

ramp_employment <- ramp_employment %>% mutate(key_worker = case_when(
  employ.government_work_key_workers == 1 | employ.government_work_key_workers.1 == 1 ~ "Government defined key worker",
  employ.government_work_key_workers == 0 | employ.government_work_key_workers.1 == 0 ~ "Not a key worker"))
table(ramp_employment$key_worker)

ramp_employment <- ramp_employment[c(1,30,31,32)]

ramp_dem <- left_join(ramp_dem, ramp_employment, by = "ID")

ramp_mhd <- readRDS(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/ramp/clinical/mhd_ramp_clean.rds")

ramp_mhd <- ramp_mhd %>% 
  mutate(anxiety_disorders = mhd.anxiety_disorders_cop) %>%
  mutate(depressive_disorders = mhd.depressive_disorders_cop) %>%
  mutate(eating_disorders = mhd.eating_disorders_cop) %>%
  mutate(obsessive_compulsive_disorders = mhd.obsessive_compulsive_disorders_cop) %>%
  mutate(psychotic_disorders = mhd.psychotic_disorders_cop) %>%
  mutate(personality_disorders = case_when(
    mhd.personality_cluster_a_cop_numeric == 1 | mhd.personality_cluster_b_cop_numeric == 1 | mhd.personality_cluster_c_cop_numeric == 1 ~ "Personality disorder",
    TRUE ~ "No personality disorder"
  )) %>%
  mutate(bipolar_disorders = case_when(
    mhd.bipolar_disorder_cop_numeric == 1 | mhd.bipolar_disorder_current_past_cop_numeric == 1 ~ "Bipolar disorder",
    TRUE ~ "No bipolar disorder"
  )) %>%
  mutate(ptsd_disorders = case_when(
    mhd.ptsd_cop_numeric == 1 | mhd.ptsd_current_past_cop_numeric == 1 ~ "PTSD",
    TRUE ~ "No PTSD"
  )) %>%
  mutate(autism_disorders = mhd.autism_spectrum_disorder_cop)

ramp_mhd <- ramp_mhd[c(1,117:125)]

ramp_dem <- left_join(ramp_dem, ramp_mhd, by = "ID")

```




```{r }

phq_mplus <- read_csv(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/phq/phq.mplus_hash_id_link_file.csv",show_col_types = FALSE)

gad_mplus <- read_csv(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/gad/gad.mplus_hash_id_link_file.csv",show_col_types = FALSE)

masq_mplus <- read_csv(file = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean/masq/masq.mplus_hash_id_link_file.csv",show_col_types = FALSE)

```

### Select relevant columns and identify factor levels

#### COPING
```{r COPING select columns and recode}
all_predictors <- fourcohort %>%
  select(ID,
         age_category_numeric,
         Gender_numeric,
         Ethnicity_numeric,
         anxiety_disorders_numeric,
         depressive_disorders_numeric,
         eating_disorders_numeric,
         obsessive_compulsive_disorders_numeric,
         psychotic_disorders_numeric,
         mhd.personality_disorder_numeric,
         mhd.bipolar_disorder_numeric,
         mhd.ptsd_numeric,
         mhd.asd_numeric,
         key_worker_numeric,
         employment_prior_covid_numeric,
         employment_change_numeric) %>%
  mutate(age = recode_factor(age_category_numeric,
                            "1" = "16-18",
                            "2" = "19-25",
                            "3" = "26-35",
                            "4" = "36-45",
                            "5" = "46-55",
                            "6" = "56-65",
                            "7" = "66-70",
                            "8" = "71-75",
                            "9" = "76+",
                            "10" = "76+",
                            "11" = "76+",
                            "12" = "76+",
                            "13"= "76+"),
       gender = recode_factor(Gender_numeric,
                            "0" = "Male",
                            "1" = "Female",
                            "2" = "Non-binary/Prefer to self define",
                            "3" = "Non-binary/Prefer to self define"),
         ethnicity = recode_factor(Ethnicity_numeric,
                                   "1" = "White",
                                   "2" = "Mixed or multiple ethnicity",
                                   "3" = "Asian",
                                   "4" = "Black",
                                   "5" = "Arab",
                                   "6" = "Other"
                                   ),
         ethnicity_clustered = recode_factor(Ethnicity_numeric,
                                   "1" = "White",
                                   "2" = "Minoritised ethnic community",
                                   "3" = "Minoritised ethnic community",
                                   "4" = "Minoritised ethnic community",
                                   "5" = "Minoritised ethnic community",
                                   "6" = "Minoritised ethnic community"
                                   ),
         anxiety_disorders =  recode(anxiety_disorders_numeric,
                                     "0" = "No anxiety disorder",
                                     "1" = "Anxiety disorder"),
         depressive_disorders =  recode(depressive_disorders_numeric,
                                     "0" = "No depressive disorder",
                                     "1" = "Depressive disorder"),
         eating_disorders =  recode(eating_disorders_numeric,
                                     "0" = "No eating disorder",
                                     "1" = "Eating disorder"),
         obsessive_compulsive_disorders =  recode(obsessive_compulsive_disorders_numeric,
                                     "0" = "No obsessive compulsive disorder",
                                     "1" = "Obsessive compulsive disorder"),
         psychotic_disorders =  recode(psychotic_disorders_numeric,
                                     "0" = "No psychotic disorder",
                                     "1" = "Psychotic disorder"),
         personality_disorders =  recode(mhd.personality_disorder_numeric,
                                     "0" = "No personality disorder",
                                     "1" = "Personality disorder"),
         bipolar_disorders =  recode(mhd.bipolar_disorder_numeric,
                                     "0" = "No bipolar disorder",
                                     "1" = "Bipolar disorder"),
        ptsd_disorders = recode(mhd.ptsd_numeric,
                                "0" = "No PTSD",
                                "1" = "PTSD"),
       autism_disorders = recode(mhd.asd_numeric,
                                 "0" = "No autism spectrum disorder",
                                 "1" = "Autism spectrum disorder"),
         key_worker =  recode(key_worker_numeric,
                                     "0" = "Not a key worker",
                                     "1" = "Government defined key worker"),
         employment_prior_covid =  recode(employment_prior_covid_numeric,
                                     "1" = "Employed",
                                     "2" = "Unemployed",
                                     "3"="Student",
                                     "4"="Retired"),
         employment_change =  recode(employment_change_numeric,
                                     "1" = "Decreased employment",
                                     "2" = "Increased employment",
                                     "3"="Furloughed",
                                     "4"="Employment not changed")) %>%
  select(ID,
         age,
         gender,
         ethnicity,
         ethnicity_clustered,
         employment_prior_covid,
         employment_change,
         key_worker,
         anxiety_disorders,
         depressive_disorders,
         eating_disorders,
         obsessive_compulsive_disorders,
         psychotic_disorders,
         personality_disorders,
         bipolar_disorders,
         ptsd_disorders,
         autism_disorders)

all_predictors$sample <- "COPING"

all_predictors <- all_predictors[, c("ID", "sample","age","gender","ethnicity","ethnicity_clustered", "employment_prior_covid","employment_change","key_worker", "anxiety_disorders", "depressive_disorders", "eating_disorders",   "obsessive_compulsive_disorders", "psychotic_disorders", "personality_disorders", "bipolar_disorders", "ptsd_disorders",  "autism_disorders" )]

      
```

#### RAMP
```{r ramp rename columns}
table(ramp_dem$dem.age_category)

names(ramp_dem)[3] <- "age"
ramp_dem <- ramp_dem %>%
  mutate(age = case_when(
    age == "16-18" ~ "16-18",
    age == "19-25" ~ "19-25",
    age == "26-35" ~ "26-35",
    age == "36-45" ~ "36-45",
    age == "46-55" ~ "46-55",
    age == "56-65" ~ "56-65",
    age == "66-70" ~ "66-70",
    age == "71-75" ~ "71-75",
    age == "76-80" | age == "81-85" | age == "86-90" | age == "91-100"| age == "100+" ~ "76+"
  )) 
table(ramp_dem$age)

table(ramp_dem$dem.which_gender_do_you_identify_with_cop)
names(ramp_dem)[4] <- "gender"
ramp_dem <- ramp_dem %>%
  mutate(gender = case_when(
    gender == "Female" ~ "Female",
    gender == "Male" ~ "Male",
    gender == "Non-binary" | gender == "Self-define" ~ "Non-binary/Prefer to self define"))
table(ramp_dem$gender)

table(ramp_dem$dem.what_is_your_ethnic_origin_cop)
names(ramp_dem)[5] <- "ethnicity"

ramp_dem <- ramp_dem %>%
  mutate(ethnicity = case_when(
    ethnicity == "White, white European or Caucasian" ~ "White",
    ethnicity == "Asian or Asian British (including Chinese)" ~ "Asian",
    ethnicity == "Mixed or multiple ethnic origins" ~ "Mixed or multiple ethnicity",
    ethnicity == "Other ethnic group (please specify)" ~ "Other",
    ethnicity == "Arab" ~ "Arab",
    ethnicity == "Black or Black British" ~ "Black")) %>%
  mutate(ethnicity_clustered = case_when(
    ethnicity == "White" ~ "White",
    ethnicity == "Mixed or multiple ethnicity" | ethnicity == "Asian" | ethnicity == "Black" | ethnicity == "Arab" | ethnicity == "Other" ~ "Minoritised ethnic community"  ))
table(ramp_dem$ethnicity)

ramp_dem <- ramp_dem[, c("ID", "sample","age","gender","ethnicity","ethnicity_clustered", "employment_prior_covid","employment_change","key_worker", "anxiety_disorders", "depressive_disorders", "eating_disorders",   "obsessive_compulsive_disorders", "psychotic_disorders", "personality_disorders", "bipolar_disorders", "ptsd_disorders",  "autism_disorders" )]
     

combine_predictors <- rbind(all_predictors, ramp_dem)
```


# left join predictors onto Mplus ID list 

This will retain only individuals in the trajectory analysis, in the correct order for merging with the Mplus output data.

```{r read in mplus id data and merge}

phq_merged <- left_join(phq_mplus, combine_predictors)
gad_merged <- left_join(gad_mplus, combine_predictors)
masq_merged <- left_join(masq_mplus, combine_predictors)


```

## save mplus linked predictor files

```{r save mplus files}
clean_data = "/Users/katherineyoung/Documents/RAMP-COPING-longitudinal_common/data_clean"

saveRDS(phq_merged, file.path(clean_data, "/phq/phq.mplus_matched_predictors.rds"))
saveRDS(gad_merged, file.path(clean_data, "/gad/gad.mplus_matched_predictors.rds"))
saveRDS(masq_merged, file.path(clean_data, "/masq/masq.mplus_matched_predictors.rds"))

```

# summary tables for predictors (for those who have data for at least one predictor)

This will represent our true sample. Others may have completed cosnent or one item, but not actually completed any questionnaires. Represents the actual sample for our trajectory analyses too as people who have no data on any predictor will not have any data for any wave of gad, phq or masq. 

First drop everyone who is NA on all variables

```{r drop all NAS}

phq_merged <- phq_merged[!apply(is.na(phq_merged[,3:16]), 1, all),]
```

```{r make kable table age phq}

age_table <-  table(phq_merged$age, useNA = "ifany") %>%
  kable(caption = "Frequency table for age in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

age_table 
```

```{r make kable table gender}

gender_table <-  table(phq_merged$gender, useNA = "ifany") %>%
  kable(caption = "Frequency table for gender in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

gender_table 
```


```{r make kable table ethnicity}

ethnicity_table <-  table(phq_merged$ethnicity) %>%
  kable(caption = "Frequency table for ethnicity in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

ethnicity_table 
```

```{r make kable table ethnicity clustered}

ethnicity_cluster_table <-  table(phq_merged$ethnicity_clustered, useNA = "ifany") %>%
  kable(caption = "Frequency table for ethnicity clustered by minoritsed ethnic groups in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

ethnicity_cluster_table 
```


```{r make kable table employment prior covid}

employment_prior_table <-  table(phq_merged$employment_prior_covid, useNA = "ifany") %>%
  kable(caption = "Frequency table for employment prior to covid in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)



employment_prior_table 
```


```{r make kable table employment change covid}
employment_change_table <-  table(phq_merged$employment_change, useNA = "ifany") %>%
  kable(caption = "Frequency table for employment change due to covid in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

employment_change_table 
```

```{r make kable table employment keyworker}

key_worker_table <-  table(phq_merged$key_worker, useNA = "ifany") %>%
  kable(caption = "Frequency table for keyworker status to covid in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

key_worker_table 
```


```{r make kable table anxiety disorders}

anxiety_disorder_table <-  table(phq_merged$anxiety_disorders, useNA = "ifany") %>%
  kable(caption = "Frequency table for anxiety disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

anxiety_disorder_table 
```


```{r}
depressive_disorder_table <-  table(phq_merged$depressive_disorders, useNA = "ifany") %>%
  kable(caption = "Frequency table for depressive disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

depressive_disorder_table 
```


```{r}
eating_disorder_table <-  table(phq_merged$eating_disorders) %>%
  kable(caption = "Frequency table for eating disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

eating_disorder_table 
```


```{r}
OCD_table <-  table(phq_merged$obsessive_compulsive_disorders) %>%
  kable(caption = "Frequency table for obsessive compulsive disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

OCD_table 
```



```{r}
 psychotic_disorders_table <-  table(phq_merged$psychotic_disorders) %>%
  kable(caption = "Frequency table for psychotic disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

psychotic_disorders_table 
```

```{r}
bipolar_disorders_table <-  table(phq_merged$bipolar_disorders) %>%
  kable(caption = "Frequency table for bipolar disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

bipolar_disorders_table 
```

```{r}
 personality_disorders_table <-  table(phq_merged$personality_disorders) %>%
  kable(caption = "Frequency table for personality disorders in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

personality_disorders_table 
```

```{r}
 ptsd_table <-  table(phq_merged$ptsd_disorders) %>%
  kable(caption = "Frequency table for PTSD in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

ptsd_table 
```

```{r}
 asd_table <-  table(phq_merged$autism_disorders) %>%
  kable(caption = "Frequency table for ASD in combined cohort (maximum n)",
        booktabs = T,
        digits = 2) %>%
  kable_styling() %>%
  row_spec(0,bold=T,background = "grey",font_size = 16)

asd_table 
```

### Save summary tables

```{r save kable}

save_file_age <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/age_summary_table.html")
save_file_gender <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/gender_summary_table.html")
save_file_employment_prepandemic <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/employment_prepandemic_summary_table.html")
save_file_employment_change <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/employment_change_summary_table.html")
save_file_ethnicity <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/ethnicity_summary_table.html")
save_file_ethnicity_clustered <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/ethnicity_clustered_summary_table.html")
save_file_keyworker <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/keyworker_summary_table.html")
save_file_anxiety <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/anxiety_disorder_summary_table.html")
save_file_depression <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/depressive_disorder_summary_table.html")
save_file_ocd <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/OCD_summary_table.html")
save_file_psychosis <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/psychosis_summary_table.html")
save_file_bipolar <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/bipolar_disorder_summary_table.html")
save_file_personality <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/personality_disorder_summary_table.html")
save_file_ed <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/eating_disorder_summary_table.html")
save_file_ptsd <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/ptsd_summary_table.html")
save_file_asd <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/asd_summary_table.html")



save_kable(age_table ,
           save_file_age)

save_kable(gender_table ,
           save_file_gender)

save_kable(employment_change_table ,
           save_file_employment_change)

save_kable(employment_prior_table ,
           save_file_employment_prepandemic)

save_kable(ethnicity_table ,
           save_file_ethnicity)

save_kable(ethnicity_cluster_table,
           save_file_ethnicity_clustered)

save_kable(key_worker_table ,
           save_file_keyworker)

save_kable(anxiety_disorder_table ,
           save_file_anxiety)

save_kable(depressive_disorder_table,
           save_file_depression)

save_kable(OCD_table ,
           save_file_ocd)

save_kable(psychotic_disorders_table,
           save_file_psychosis)

save_kable(bipolar_disorders_table ,
           save_file_bipolar)

save_kable(personality_disorders_table ,
           save_file_personality)

save_kable(eating_disorder_table ,
           save_file_ed)

save_kable(ptsd_table ,
           save_file_ptsd)

save_kable(asd_table ,
           save_file_asd)

```

#combined summary table for paper
```{r combined summary table for paper}

if(!require(expss)){
  library(expss)
  install.packages("expss",repos = "http://cran.us.r-project.org")
}


#remove_column(age_table,4:6)

#dd %>% tab_cells("Sex" = sex, "Marital status" = marital, "Age" = age_group) %>% 
#    tab_cols(total(label = "N")) %>% 
#    tab_pivot() %>% 
#    tab_stat_cases(total_row_position = "none") %>% 
#    split_columns() %>% 
#    kable()

combined_table <- phq_merged %>% tab_cells("Age" = age, "Gender" = gender,"Ethnicity" = ethnicity, "Ethnicity - clustered" = ethnicity_clustered, "Change in employment" = employment_change, "Pre-pandemic employment" = employment_prior_covid, "Keyworker status" = key_worker, "Depressive disorder diagnosis" = depressive_disorders, "Anxiety disorder diagnosis" = anxiety_disorders, "OCD diagnosis" = obsessive_compulsive_disorders, "Psychotic disorder diagnosis" = psychotic_disorders, "Bipolar disorder diagnosis" = bipolar_disorders,"Eating disorder diagnosis" = eating_disorders, "Personality disorder diagnosis" = personality_disorders, "PTSD diagnosis" = ptsd_disorders, "ASD diagnosis" = autism_disorders) %>% 
    tab_cols(total(label = "N")) %>% 
    tab_stat_cases(total_row_position = "none") %>% 
    tab_pivot() %>% 
    kable()


save_file_combi <- file.path(dirname(dirname(getwd())),"output/baseline_predictor_summary_tables/combi_table.html")

save_kable(combined_table ,
           save_file_combi)

combined_table

```
##by cohort - combined summary table for paper
```{r by cohort combined summary table for paper}

coping_table <- phq_merged %>% filter(sample == "COPING") %>% tab_cells("Age" = age, "Gender" = gender,"Ethnicity" = ethnicity, "Ethnicity - clustered" = ethnicity_clustered, "Change in employment" = employment_change, "Pre-pandemic employment" = employment_prior_covid, "Keyworker status" = key_worker, "Depressive disorder diagnosis" = depressive_disorders, "Anxiety disorder diagnosis" = anxiety_disorders, "OCD diagnosis" = obsessive_compulsive_disorders, "Psychotic disorder diagnosis" = psychotic_disorders, "Bipolar disorder diagnosis" = bipolar_disorders,"Eating disorder diagnosis" = eating_disorders, "Personality disorder diagnosis" = personality_disorders, "PTSD diagnosis" = ptsd_disorders, "ASD diagnosis" = autism_disorders) %>% 
    tab_cols(total(label = "N")) %>% 
    tab_stat_cases(total_row_position = "none") %>% 
    tab_pivot() %>% 
    kable()

coping_table

ramp_table <- phq_merged %>% filter(sample == "RAMP") %>% tab_cells("Age" = age, "Gender" = gender,"Ethnicity" = ethnicity, "Ethnicity - clustered" = ethnicity_clustered, "Change in employment" = employment_change, "Pre-pandemic employment" = employment_prior_covid, "Keyworker status" = key_worker, "Depressive disorder diagnosis" = depressive_disorders, "Anxiety disorder diagnosis" = anxiety_disorders, "OCD diagnosis" = obsessive_compulsive_disorders, "Psychotic disorder diagnosis" = psychotic_disorders, "Bipolar disorder diagnosis" = bipolar_disorders,"Eating disorder diagnosis" = eating_disorders, "Personality disorder diagnosis" = personality_disorders, "PTSD diagnosis" = ptsd_disorders, "ASD diagnosis" = autism_disorders) %>% 
    tab_cols(total(label = "N")) %>% 
    tab_stat_cases(total_row_position = "none") %>% 
    tab_pivot() %>% 
    kable()

ramp_table
```
